---
title: "EDRM 718 Final Exam"
author: "Ningqiao Li"
date: "May 4th, 2020"
output: html_notebook
---


This is final exam report of EDRM 718 created by Ningqiao Li

***

### 1. Tidy the data and merge them into one csv file.
#### 1.1 Tidy the Job Prep Demographics
##### 1.1.1 Install package first, and manually save the data file as csv format.
```{r include=FALSE}
library(tidyverse)
library(here)
```


##### 1.1.2 Tidy the data then
```{r}
demographics <- read_csv("Job Prep Demographics.csv", skip = 4,
                         col_types = cols(HUMANITIES = col_double(),
                                          X2 = col_double(),
                                          X3 = col_double(),
                                          X4 = col_factor()))
View(demographics)
demographics$X5 <- NULL
```


##### 1.1.3 Split the dataset into five subsets.
```{r}
demographics1 <- demographics %>%
  filter(between(row_number(), 3, 140)) %>%
  mutate(discipline = "Humanities")

demographics2 <- demographics %>%
  filter(between(row_number(), 146, 298)) %>%
  mutate(discipline = "SocialSci")

demographics3 <- demographics %>%
  filter(between(row_number(), 304, 458)) %>%
  mutate(discipline = "NaturalSci")

demographics4 <- demographics %>%
  filter(between(row_number(), 464, 612)) %>%
  mutate(discipline = "FormalSci")

demographics5 <- demographics %>%
  filter(between(row_number(), 618, 761)) %>%
  mutate(discipline = "AppliedSci") 
```


##### 1.1.4 Merge the five tidy subsets. 
```{r}
newdemo <- rbind(demographics1, demographics2, demographics3, demographics4, demographics5)  
colnames(newdemo) = c("ID", "SchoolCode", "Age", "Gender", "Discipline") 
  
(newdemo <- newdemo %>%
  arrange(ID))
```


#### 1.2 Tidy the Job Prep GPA Data
```{r}
GPA <- read_csv("Job Prep GPA Data.csv", skip = 5,
                col_names = c("ID", "SchoolCode", "Age", "Gender", 
                             "CreditsS1", "GPAS1", "CreditsS2", "GPAS2"))
(GPA <- GPA %>%
    arrange(ID))
```


#### 1.3 Join the two datasets
```{r}
merge1 <- GPA %>%
  left_join(newdemo, c("ID" = "ID", "SchoolCode" = "SchoolCode", "Age" = "Age", "Gender" = "Gender"))
```


***


### 2. Compute the average GPA.
#### 2.1 Visually display the consistency of first and second semester GPA
```{r}
# gather data first
(merge2 <- merge1 %>%
  gather(GPAS1, GPAS2,
         key = Semester,
         value = GPA))

# boxplot
merge2 %>%
  ggplot(aes(Semester, GPA)) +
  geom_boxplot() +
  labs(title = "Table 1. Comparison between GPA of two semesters")

# descriptive statistics
tapply(merge2$GPA, merge2$Semester, summary)

# independent t test
t.test(merge2$GPA ~ merge2$Semester, pair = TRUE)

# The result of descriptive statistics shows that the two groups have similar mean and median value,
# the t test confirms that there is no significant difference between GPA in semester 1 and semester 2. 
# Based on the boxplot, descriptive statistics and t test, we can conclude that the GAP in two semesters are consistent.
```


#### 2.2 Exclude the students with ... ten percent largest residuals
```{r}
# spread data
(merge1 <- merge2 %>%
  spread(key = Semester, value = GPA))


# using GPAS1 to predict GPAS2 to obtain the residuals
merge1 %>%
  ggplot(aes(x = GPAS1, y = GPAS2)) +
  geom_point() +
  labs(title = "Table 2. Relationship between GPAS1 and GPAS2") +
  stat_smooth(method = "lm", se = FALSE)

# employ the function to obtain residuals and exclude the students with the 10% largest value of residual, 
# and then generate a new dataset.

dataset_exclude <- obtain_residual(merge1, merge1$GPAS1, merge1$GPAS2, 0.1)
dataset_exclude

# highlight those students with large value in plot.
scatter_plot <- plot(dataset_exclude$GPAS1 ~ dataset_exclude$GPAS2,
                     main = "Table 3. Regression predicting", 
                     xlab = "GPA of the first semester", ylab="GPA of the second semester",
                     col = ifelse(dataset_exclude$result_res,"black","red"))   # no red point, why?

abline(lm(GPAS2 ~ GPAS1, data = dataset_exclude), col = "red")
```


#### 2.3 weighted average GPA
```{r}
dataset_exclude$result_res <- NULL
dataset_exclude$result_res_abs <- NULL

(dataset_exclude <- dataset_exclude %>%
    mutate(AVEGPA = (GPAS1 * CreditsS1 + GPAS2 * CreditsS2)/(CreditsS1 + CreditsS2)))
```


***


### 3. Merge all the three data set
#### 3.1 Tidy the Job Prep Survey Results
```{r}
survey <- read_csv("Job Prep Survey Results.csv", skip = 3)
# employ the sourced new_survey_data function to account for reverse score.
source(here("Functions", "Reverse Survey Score.R"))
newsurvey <- new_survey_data(survey, 6, c(5, 10, 13, 14))  
newsurvey
newsurvey[, 2:15] <- NULL
newsurvey
```

#### 3.2 Merge the survey data to the merge1 dataset, thus all three original datasets are tidy and merged into one dataset.
```{r}
(new_dataset <- newsurvey %>%
  left_join(dataset_exclude, c("ID" = "ID")) %>%
    arrange(ID))

# rename the Sum to as Optimism
new_dataset <- new_dataset %>%
  rename(Optimism = Sum)
new_dataset

# check for NA data
colSums(is.na(new_dataset))
(new_dataset <- new_dataset %>%
  drop_na(SchoolCode))

```



***


### 4. Response to research Q1, 
### Is senior-year grade point average (GPA) related to the degree of optimism about future employment?

```{r}
# use correlation test whether average GPA is correlated with the degree of optimism.
cor.test(new_dataset$AVEGPA, new_dataset$Optimism)

# visualize the correlation
library(ggpubr)
ggscatter(new_dataset,
          x = "AVEGPA", y = "Optimism",
          add = "reg.line", conf.int = TRUE,
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Average GPA", ylab = "Optimism of graduation",
          title = "Table 4. Correlation between average GPA and graduation optimism")

# Based on the scatterplot and the p value of pearson correlation, 
# there is no significant correlation relationship between average GPA and the degree of optimism about future employment.

MODEL <- lm(new_dataset$Optimism ~ new_dataset$AVEGPA)
summary(MODEL)

# In addition, from the result of sample linear regression, the insignificant coeficient of explanatory variable also 
# confirms the previous conclusion.

```


***


### 5. Response to research Q2,
### Is the relationship of GPA and optimism consistent or different for different categories of disciplines?

```{r}

# plot a graph first to see the relationship between GPA and optimism by different categories.
new_dataset %>%
  ggplot(aes(x = AVEGPA, y = Optimism)) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE) +
  facet_wrap(~ Discipline)
# from the graph we can see that the relationship presents slightly positive trend in Applied Science, Humanities, 
# Social Science, and a slightly negative relationship in Formal Science and Natural Science.However, we still do 
# the correlation test to confirm the relationship.


# test the correlation among different categoires.
new_dataset %>%
group_by(Discipline) %>%
summarize(cor(AVEGPA, Optimism))
# the correlation test confirms the relationship between average GPA and Optimism is positive in applied science, social science and humanities, with the largest correlation coefficients in humanities. Besides, the relationship between average GPA and Optimism is negative in formal science and natural science. 

```


***


### 6. Response to research Q3,
### Is the level of optimism different for different categories of disciplines?

```{r}
# descriptive statistics
tapply(new_dataset$Optimism, new_dataset$Discipline, summary)

# omnibus hypotheses test by ANOVA
anova(lm(new_dataset$Optimism ~ new_dataset$Discipline))
# the result shows that at 5% significance level, difference doesn't exist among diffent categories of disciplines regarding level of optimism with 95% confidence. 

# next compute Tukey HSD to perform multiple pairwise-comparison between the means of groups
# method 1
aovresult <- aov(Optimism ~ Discipline, data = new_dataset)
TukeyHSD(aovresult)

# method 2
pairwise.t.test(new_dataset$Optimism, new_dataset$Discipline,
                p.adjust.method = "BH")
# the result of two methods confirm that at 95 confidence level, there is no significant difference between the pairs.
# except the mean difference between social science and applied science is different at 10% significance level.
# Meanwhile, the 95% confidence intervel is construct as the TukeyHSD function shows.
```






